OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-04-08 paper/paper.Rmd
Improve split dataset results

- we wanted to use x% of data as ref, weren't sure what size to use and strategy for selecting.
- important result: it works!

@@

**TODO; ref weight simple only here. First we wanted to know; does this strategy perform as well as de novo with 50% split?**
OTU quality from the split dataset strategy with OptiFit was highly similar to
that from _de novo_ clustering the whole dataset with OptiClust regardless of
mode (`r mcc_opticlust_vs_fit_split_simple`% difference in median MCC).
OTU quality was remarkably stable across reference fraction sizes within
parameter sets.
For example, splitting the human dataset 100 times yielded a coefficient of
variation of `r cv_fit_split_mcc_human` for the MCC score.
In terms of runtime, closed and open reference OptiFit performed faster than
OptiClust on whole datasets by `r sec_closed_fit_split_vs_clust`% and 
`r sec_open_fit_split_vs_clust` respectively.
The split dataset strategy performed `r sec_closed_fit_split_vs_db`% faster than
the database strategy in closed reference mode and `r sec_open_fit_split_vs_db`%
faster in open reference mode.
**TODO: fraction mapped.**

**Then we wanted to know; what fraction of sequences should be in the reference?**
To test the best reference size, reference sizes from 10% to 80% of the sequences were created, with the remaining sequences used for the query.

>>>>

@@

First, we tested whether OptiFit performed as well as _de novo_ clustering when
using the split dataset strategy with half of the sequences selected for
the reference by a simple random sample.
OTU quality was highly similar to that from OptiClust regardless of
mode (`r mcc_opticlust_vs_fit_split_simple`% difference in median MCC).
In closed reference mode, OptiFit was able to fit `r frac_fit_split`% of query
sequences to reference OTUs with the split strategy, a great improvement over
the average `r frac_fit_gg`% of sequences fit to the greengenes database.
In terms of runtime, closed and open reference OptiFit performed faster than
OptiClust on whole datasets by `r sec_closed_fit_split_vs_clust`% and 
`r sec_open_fit_split_vs_clust` respectively.
The split dataset strategy also performed `r sec_closed_fit_split_vs_db`% faster
than the database strategy in closed reference mode and 
`r sec_open_fit_split_vs_db`% faster in open reference mode.
Thus, reference clustering with the split dataset strategy creates as high
quality OTUs as _de novo_ clustering yet at a faster run time, and fits far more
query sequences than the database strategy.

**Then we wanted to know; what fraction of sequences should be in the reference?**
To test the best reference size, reference sizes from 10% to 80% of the
sequences were created, with the remaining sequences used for the query.
OTU quality was remarkably stable across reference fraction sizes.
For example, splitting the human dataset 100 times yielded a coefficient of
variation of `r cv_fit_split_mcc_human` for the MCC score across all fractions. **TODO: revisit how to report this**
